---
title: "Assignment 1 _Meinhardt"
author: "Julia Meinhardt"
date: "9/6/2020"
output: html_document
---

# Welcome

Hello,  world. (This is markdown text)

## It's me. 

Here I am. 

```{r}

```

That is a code chunk ^
Knit it together when you're ready!
When you knit, you'll create an html file. 
When you create a markdown, you create a .rmd.
All of this is living in the .rproj repo! 

When you're ready to push it to github, go to the git tab. Check the boxes beside the files you want to push to github. Click commit. Leave a little message to track what you changed in the repo. Commit. Then PUSH. 

To load libaries, use the following:

   {r load libraries}
library(tidycensus)
library(tidyverse)

```{r load libraries}
library(tidycensus)
library(tidyverse)

```

CTRL, SHIFT, ENTER to run a line of code or run button ahead 

```{r load person-level microdata variables}
person_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "person", year == 2018, survey == "acs1")
```

```{r load household-level microdata variables }
hh_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```


Loading Data

Variables
1. SCHL - Education Level (categorical)
2. SEX (categorical)
3. WAGP - Wage/salary for one year (continuous)
4. WKHP - Hours of work (continuous)
5. WMNP - Travel time to work (continuous)

Since every person is in a household, you can also include household data for a person-level analysis. (i.e. tenure? how long they've lived in their current location)

```{r load person level data, results = 'hide'}
person_data <- get_pums(variables = c("SCHL",
                                      "AGEP",
                                      "SEX", 
                                       "WAGP", 
                                      "WKHP", 
                                      "JWMNP"),
                        state = "06", 
                        geography = "county",
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE)
  filter(JWMNP > 0, AGEP > 18) 
  select(SCHL_label, SEX_label, WAGP, WKHP, JWMNP)
```


```{r show person-level data}
person_data
```



CODE GRAVEYARD: Was trying to get data on the female population of different rural counties across NorCal:

Problems with loading variables in chart. 

1. Total Female Population: B01001_026 
2. Insurance Coverage: B27001_030
3. Travel Time: B08012_02
4. Marital Status: B12501_007
5. Poverty Level : B17001_017
6. Education Level: B14002_026
7. Earnings : B20001_023

```{r}
pop_vars <- c(total_female = "B01001_001")
            

tractModoc2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Modoc",
              year = 2018, 
              survey = "acs5")

tractModoc2018$county <- rep("Modoc", times = length(tractModoc2018$total_femaleE))

tractLassen2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Lassen",
              year = 2018, 
              survey = "acs5")

tractLassen2018$county <- rep("Lassen", times = length(tractLassen2018$total_femaleE))

tractPlumas2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Plumas",
              year = 2018, 
              survey = "acs5")

tractPlumas2018$county <- rep("Plumas", times = length(tractPlumas2018$total_femaleE))

tractShasta2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Shasta",
              year = 2018, 
              survey = "acs5")

tractShasta2018$county <- rep("Shasta", times = length(tractShasta2018$total_femaleE))

tractTrinity2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Trinity",
              year = 2018, 
              survey = "acs5")

tractTrinity2018$county <- rep("Trinity", times = length(tractTrinity2018$total_femaleE))

tractSiskiyou2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Siskiyou",
              year = 2018, 
              survey = "acs5")

tractSiskiyou2018$county <- rep("Siskiyou", times = length(tractSiskiyou2018$total_femaleE))

tractDelNorte2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Del Norte",
              year = 2018, 
              survey = "acs5")

tractDelNorte2018$county <- rep("Del Norte", times = length(tractDelNorte2018$total_femaleE))

```



```{r mergecounties}
jeffersonCA2018 <- rbind(tractModoc2018, tractLassen2018, tractPlumas2018, tractShasta2018, tractTrinity2018, tractSiskiyou2018, tractDelNorte2018)
```



