---
title: 'Assignment 7: Raster Interpolation'
author: "Julia Meinhardt"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
```

Identify an area-level data set for a single variable (might be at the neighborhood level, zip-code level, or county-level, for example) that includes at least ten different zones. Show the variable in that variable by zone (as a chloropleth map), as points (using area centroids), and as a continuous surface (interpolating between centroid points. Your submission should include a discussion of which of these three methods is (a) most informative, (b) most interesting, (c) most appropriate to the data, and (d) best.

```{r}
counties <- st_read("https://public.opendatasoft.com/explore/dataset/us-county-boundaries/download/?format=geojson&timezone=America/Los_Angeles&lang=en", quiet = TRUE) %>%
 dplyr::select(name) 

leaflet(counties) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillColor = "yellow", 
                                                  fillOpacity = 1),
                                                  label = ~name, 
                                                  weight = 1) 


```
```{r}

countyincome <- read.csv(file = "Income.csv")

```

```{r}
countyincome <- dplyr::rename(countyincome, name = County) 
```

```{r}
countypoverty <- dplyr::left_join(countyincome, counties, by = "name")
```




